---
title: "Chrononormativity Canadian 2020 Analysis"
author: "James Steur & Aleks Ksiazkiewicz"
date: "01/24/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r Load Packages & Data, echo=FALSE, include=FALSE, results=FALSE}
#Load in Packages & Data
library(haven)
CES_2021_Alek_Module_weighted_final_2 <- read_dta("~/Desktop/CanadianAnalysis/CES 2021_Alek_Module_weighted_final_2.dta")
canada <- CES_2021_Alek_Module_weighted_final_2
library(readr)
library('devtools')
library('mirt')
#install_github('philchalmers/mirt')
library(dplyr)
packageVersion("dplyr")
library("ltm")
library('mvtnorm')
library(haven)
library('stargazer')
library('ggpubr')
library(dplyr)
library(car)
library(Hmisc)
library(effectsize)
library(haven)
library('naniar')
```

```{r, Data Cleaning, echo = FALSE, include=FALSE, results=FALSE}
#Data Cleaning
canada <- replace(canada, canada == -99, NA) #Replaces with NA values

#Recode Chrononormativity & Sum Variable
chrononormativity <- subset(canada,
  select=c(cps21_Alex_Q4_1:cps21_Alex_Q4_17))
chrononormativity[chrononormativity==2] <- 0 
#There are proper or improper times of day to do this activity (1)
#People should do this activity at whatever time of day suits them (0)
canada$sum_normativity <- rowSums(chrononormativity)
table(canada$sum_normativity)

#Code Chronotype

sleep1 <- subset(canada, select = c(cps21_Alek_Q1:cps21_Alek_Q3))
sleep1$cps21_Alek_Q1 <- sleep1$cps21_Alek_Q1 + 20
sleep1$cps21_Alek_Q3 <- sleep1$cps21_Alek_Q3/30
sleep1$GenDuration <- (sleep1$cps21_Alek_Q1 - sleep1$cps21_Alek_Q2)

sleep1$GenDuration <- with(sleep1, 
  ifelse(GenDuration >= 48, GenDuration - 48, 
  ifelse(GenDuration < 0, GenDuration + 48,
  GenDuration)))

sleep1$GenDuration[sleep1$GenDuration == 0] <- NA

sleep1$cps21_Alek_Q3 <- with(sleep1,
  ifelse(GenDuration - cps21_Alek_Q3 <= 0, NA, cps21_Alek_Q3))

#Exclude Outliers on sleep1/wake times. 
#UIL 303. Before 7PM = 1 or 2. After 6AM = 25 to 48.
sleep1$cps21_Alek_Q2 <- with(sleep1,
  ifelse(cps21_Alek_Q2 <= 2, NA, 
  ifelse(cps21_Alek_Q2 >= 25, NA, 
  cps21_Alek_Q2)))

#UIL 302. Before 4AM = 21 or less. After 3PM = 43 to 48.
sleep1$cps21_Alek_Q1 <- with(sleep1,
  ifelse(cps21_Alek_Q1 <= 21, NA,
  ifelse(cps21_Alek_Q1 >= 43, NA,
  cps21_Alek_Q1)))

#Remove sleep1ing fewer than 2 hours or more than 18 hours.
sleep1$GenDuration <- with(sleep1,
  ifelse(GenDuration <= 4, NA,
  ifelse(GenDuration >= 36, NA,
  GenDuration)))

#Midpoint of free night sleep.
sleep1$GenMfs <- sleep1$cps21_Alek_Q1 - (sleep1$GenDuration*0.5)

sleep1$GenMfs <- (ifelse(sleep1$GenMfs > 48, sleep1$GenMfs - 48, sleep1$GenMfs))

sleep1$delay <- sleep1$cps21_Alek_Q1 - (sleep1$GenDuration - sleep1$cps21_Alek_Q3)/2 
sleep1$delay <- ifelse(sleep1$delay > 48, sleep1$delay-48, sleep1$delay)

#Below 13 = Before Midnight. After 13 = After Midnight. Diving by two to get hourly unit. 
sleep1$GenMfs <- (sleep1$GenMfs - 13)/2
sleep1$delay <- (sleep1$delay - 13)/2

#Use delay for analysis.
canada$chronotype_delay <- sleep1$delay

#3.5 is the median amount of midpoint sleep. 
#3.5 or below = Morning. 1.
#Higher than 3.5 = Evening. 2. 

sleep1$bi_type <- with(sleep1,
  ifelse(delay <= 3.5, 1, 
  ifelse(delay > 3.5, 2,
         delay)))

canada$bi_type <- sleep1$bi_type
canada$bi_type
```

```{r, Correlation & Regression Models, echo = FALSE, include=FALSE, results=FALSE}
#Correlation Recode
cor_frame <- subset(canada, select=c(cps21_lr_scale_bef_1, chronotype_delay, sum_normativity))
names(cor_frame)[names(cor_frame) == "sum_normativity"] <- "Chrononormativity"
names(cor_frame)[names(cor_frame) == "cps21_lr_scale_bef_1"] <- "Ideology"
names(cor_frame)[names(cor_frame) == "chronotype_delay"] <- "Chronotype"
canada$cps21_lr_scale_bef_1
cor_frame1 <- rcorr(as.matrix(cor_frame))
cor_frame1

m1 <- lm(cps21_lr_scale_bef_1 ~ sum_normativity, data = canada)
m2 <- lm(cps21_lr_scale_bef_1 ~ sum_normativity + chronotype_delay, data = canada)
m3 <- lm(cps21_lr_scale_bef_1 ~ sum_normativity + chronotype_delay + cps21_age, data = canada)
m4 <- lm(cps21_lr_scale_bef_1 ~ sum_normativity + chronotype_delay + cps21_age + cps21_genderid, data = canada)
m5 <- lm(cps21_lr_scale_bef_1 ~ sum_normativity + chronotype_delay + cps21_age + cps21_genderid + cps21_education, data = canada)
m6 <- lm(cps21_lr_scale_bef_1 ~ sum_normativity + chronotype_delay + cps21_age + cps21_genderid + cps21_education + cps21_fed_id, data = canada)


m7 <- lm(chronotype_delay ~ sum_normativity, data = canada)
m8 <- lm(chronotype_delay ~ sum_normativity + cps21_lr_scale_bef_1, data = canada)
m9 <- lm(chronotype_delay ~ sum_normativity + cps21_lr_scale_bef_1 + cps21_age, data = canada)
m10 <- lm(chronotype_delay ~ sum_normativity + cps21_lr_scale_bef_1 + cps21_age + cps21_genderid, data = canada)
m11 <- lm(chronotype_delay ~ sum_normativity + cps21_lr_scale_bef_1 + cps21_age + cps21_genderid + cps21_education, data = canada)
m12 <- lm(chronotype_delay ~ sum_normativity + cps21_lr_scale_bef_1 + cps21_age + cps21_genderid + cps21_education + cps21_fed_id, data = canada)

#The Canadian Study didn't measure race/ethnicity... so, that's that. 
#canada$cps21_age #Age
#canada$cps21_genderid #Gender
#canada$cps21_education Education

#canada$cps21_lr_scale_bef_1 #Self placement ideology.
#canada$cps21_fed_id. Partisanship. Eight Different Parties. 
table(canada$cps21_fed_id)
```

# Pearson Correlations

Ideology, chronotype, and chrononormativity don't correlate strongly with one another. Correlations, on average around -1 or 1. All p-values are significant except for chrononormativity and ideology. Significant p-values are not unexpected given the large n-sizes. 
```{r, Correlations Bivariate Regressions}
cor_frame1
```
# Bivariate Regression Models

There is not a significant p-value for the relationship between chrononormativity (IV) and ideology (DV). With controls included, chronotype, age, gender, and education are significant.

There is a significant p-value for the relationship between chrononormativity (IV) and chronotype (DV). With controls included, ideology, age, and chrononormativity are significant. 

Model 1 is a bivariage regression with chrononormativity as the IV and ideology as the DV.   
Model 7 is a bivariate regression with chrononormativity as the IV and chronotype as the DV. 
  
The other models include all of the other controls: age, income, gender, and party. Model 6 also has chronotype as a control, and model 12 has ideology as a control.  

*Variable Key*  
sum_normativity is chrononormativity  
chronotype_delay is chronotype  
cps21_lr_scale_bef_1 is ideology  

```{r, Bivariate Regressions}
summary(m1) 
summary(m6)
summary(m7)
summary(m12)
```

\newpage


# Difference in Means

Chronotype has a significant p-value in the analysis. Interestingly, it appears as though morning types were more likely to relax in the system affirming condition. 

```{r, echo=FALSE, include=FALSE, results=FALSE}
#Two Way Anova
anova1 <- subset(canada, select=c(sum_normativity, Alek_module_condition, bi_type))
anova1$Chrononormativity <- anova1$sum_normativity
anova1$Chronotype <- anova1$bi_type
anova1$Treatment <- anova1$Alek_module_condition

anova1$Treatment <- factor(anova1$Treatment, 
                  levels = c("C1_Control", "C2_System_threat", "C3_System_affirming"),
                  labels = c("Control", "Threat", "Affirming"))

anova1$Chronotype <- factor(anova1$Chronotype, 
                  levels = c(1, 2),
                  labels = c("Morning", "Evening"))

group_by(anova1, Treatment, Chronotype) %>%
  summarise(
    count = n(),
    mean = mean(Chrononormativity, na.rm = TRUE),
    sd = sd(Chrononormativity, na.rm = TRUE)
  )
```

```{r, echo=FALSE}
group_by(anova1, Treatment, Chronotype) %>%
  summarise(
    count = n(),
    mean = mean(Chrononormativity, na.rm = TRUE),
    sd = sd(Chrononormativity, na.rm = TRUE)
  )
```


\newpage
Higher levels on the y-axis for chrononormativity reflect thinking there is a proper or improper time for certain activities. Lower scores indicate thinking there is not a proper or improper time for certain activities. 
```{r, echo=FALSE, warning = FALSE}
ggboxplot(anova1, x = "Treatment", y = "Chrononormativity", color = "Chronotype",
          palette = c("#00AFBB", "#E7B800"))

two.way <- aov(Chrononormativity ~ Treatment*Chronotype, data = anova1)
summary(two.way)


hist(canada$sum_normativity,
     main = "Chrononormativity Responses",
     xlim = c(0,20),
     ylim = c(0,600),
     xlab = "Chrononormativity", 
     col = "blue")

my_anova1 <- aov(Chrononormativity ~ Treatment*Chronotype, data = anova1)
#imbalance <- Anova(my_anova, type = "III")

summary(my_anova1)
Anova(two.way, type = "III")
```
\newpage


# IRT

All items except number 16 discriminate with values over 1.

All of the items also fit the model.

The last three items are capturing the most variation in terms of people. That is, more extreme ends of the curve/middle of the curve are captured.

Test Information Curve is capturing roughly 1 standard deviation above the norm on all items. (The Item Characteristic Curves are included for all items. The first graph represents item 1, the second graph represents item 2, etc.)
```{r, IRT, echo=FALSE}
#2PL Binary IRT. #Norm_sum is the subset of the chrononormativity items. 
BinModel1 <- ltm(chrononormativity ~ z1, IRT.param = T) #2PL model. 
#summary(BinModel1) #Difficulty, Discrimination, & EFA
#coef(BinModel1, IRTpars = T) #Coefficients
coef(BinModel1) #Nicely Formatted Difficulty & Discrimination
item.fit(BinModel1)

plot(BinModel1, type = "IIC", items = 0) #Test Information Function

#Item Characteristic Curve of All Items
plot(BinModel1, type = "ICC", items = 1)
plot(BinModel1, type = "ICC", items = 2)
plot(BinModel1, type = "ICC", items = 3)
plot(BinModel1, type = "ICC", items = 4)
plot(BinModel1, type = "ICC", items = 5)
plot(BinModel1, type = "ICC", items = 6)
plot(BinModel1, type = "ICC", items = 7)
plot(BinModel1, type = "ICC", items = 8)
plot(BinModel1, type = "ICC", items = 9)
plot(BinModel1, type = "ICC", items = 10)
plot(BinModel1, type = "ICC", items = 11)
plot(BinModel1, type = "ICC", items = 12)
plot(BinModel1, type = "ICC", items = 13)
plot(BinModel1, type = "ICC", items = 14)
plot(BinModel1, type = "ICC", items = 15)
plot(BinModel1, type = "ICC", items = 16)
plot(BinModel1, type = "ICC", items = 17)

#plot(BinModel1, type = "ICC") #All items at once
#plot(BinModel1, type = "ICC", items = 3)


#Factor_Scores <- factor.scores(BinModel1) 
#See Frequency of How Often People Select

#person.fit(BinModel1)
```

```{r, Key Variables, echo = FALSE, include=FALSE, results=FALSE}
#Exploratory Analyses: Territory & Religion. 

#canada$Alek_module_condition #Treatment Condition. 1 = Control. 2 = Threat. 3 = Affirming.
#canada$cps21_Alek_Q1 # Free Days.
#canada$cps21_Alek_Q2 # Nights
#canada$cps21_Alek_Q3 #Minutes before sleep
#canada$cps21_Alex_Q4_1 #1-17 is chrononormativity scale.
#canada$cps21_age #Age
#canada$cps21_genderid #Gender
#canada$cps21_education Education

#canada$cps21_lr_scale_bef_1 #Self placement ideology.
#canada$cps21_fed_id. Partisanship. Eight Different Parties. 

#table(canada$cps21_data_quality) #Address poor quality responses in data. Include poor respondents with analysis and exclude them to see if differences arise in analysis. 
```

